Associations between Chlorophyll *a* and various Microcystin-LR Health Advisory

### 2015-09-11 manuscript/manuscript.Rmd
started draft figure layouts

who_rec_low2_cp <- condprob(xX = nla_dat$log_chla,xY = nla_dat$log_micro, 
                     xImpair = log10(mc_guides[5]), ProbComp = "gt", Exceed = "gte",
                     ci = TRUE, R = 100, xW=wgt)

mc_chla[5] <- get_chla(who_rec_low2_cp,0.5)

who_rec_med1_cp <- condprob(xX = nla_dat$log_chla,xY = nla_dat$log_micro, 
                     xImpair = log10(mc_guides[6]), ProbComp = "gt", Exceed = "gte",
                     ci = TRUE, R = 100, xW=wgt)

mc_chla[6] <- get_chla(who_rec_med1_cp,0.5)

who_rec_med2_cp <- condprob(xX = nla_dat$log_chla,xY = nla_dat$log_micro, 
                     xImpair = log10(mc_guides[7]), ProbComp = "gt", Exceed = "gte",
                     ci = TRUE, R = 100, xW=wgt)

mc_chla[7] <- get_chla(who_rec_med2_cp,0.5)
```

@@

#Methods
We evaluated associated chlorophyll *a* concentrations for an effect for each of the WHO and U.S. EPA levels (Table \ref{tab:microcystin_levels}). Lakes with higher microcystin-LR concentrations were rare. Only `r round(100*(length(nla_dat$MCYST_TL_UGL[nla_dat$MCYST_TL_UGL>=10])/nrow(nla_dat)),2)` % of lakes sampled had a concentration greater than 10.  For this analysis we focus on the microcystin concentrations that are better represented in the NLA data. These were  0.3, 1, 1.6, and 2 $\mu$g/L.

@@

(Figure \ref{fig:epa_child_cp_plot})

(Figure \ref{fig:who_drink_cp_plot})

(Figure \ref{fig:epa_adult_cp})

(Figure \ref{fig:who_rec_low1_cp})

@@

```{r epa_child_cp_plot,echo=FALSE, fig.cap="Plot of \\label{fig:epa_child_cp_plot}"}
plot_cp(epa_child_cp,x = "Log 10 Chlorophyll", y = "Probability of Exceeding EPA Child HA")
```

@@

```{r who_drink_cp_plot,echo=FALSE, fig.cap="Plot of  \\label{fig:who_drink_cp_plot}"}
plot_cp(who_drink_cp, x = "Log 10 Chlorophyll", y = "Probability of Exceeding WHO Drinking HA")
```

@@

```{r epa_adult_cp ,echo=FALSE, fig.cap="Plot of  \\label{fig:epa_adult_cp}"}
plot_cp(epa_adult_cp, x = "Log 10 Chlorophyll", y = "Probability of Exceeding EPA Adult HA")
```

```{r who_rec_low1_cp, echo=FALSE, fig.cap="Plot of  \\label{fig:who_rec_low1_cp}"}
plot_cp(who_rec_low1_cp, x = "Log 10 Chlorophyll", y = "Probability of Exceeding WHO Low Probability HA")
```

@@

```{r microcystin_levels,echo=FALSE,results='asis', fig.cap="Various suggested microcystin-LR health advisory concentrations. \\label{fig:microcystin_levels}"}
source_of_stnd <- c("WHO",
                    "WHO",
                    "WHO",
                    "WHO",
                    "WHO",
                    "U.S. EPA",
                    "U.S. EPA")
type_of_stnd <- c("Recreational",
                  "Recreational",
                  "Recreational",
                  "Recreational",
                  "Drinking",
                  "Drinking",
                  "Drinking")
concentration <- c("2-4 ug/l",
                   "10-20 ug/l",
                   "20-2000 ug/l",
                   ">2000 ug/l",
                   "1 ug/l",
                   "0.3 ug/l",
                   "1.6 ug/l")
#note <- c("Low probability of adverse health effects",
#          "Moderate probability of adverse health effects",
#          "High probability of adverse health effects, These higher #concentrations typically only seen in visible scums/algal mats",
#          "Very high probability of adverse health effects",
#          "",
#          "Children younger than school age",
#          "Individuals School age and older")
mc_conc_table <- data.frame(source_of_stnd, type_of_stnd, concentration) %>%
  arrange(type_of_stnd)
names(mc_conc_table)<-c("Source","Type","Concentration")
pandoc.table(mc_conc_table,justify = "left" ,split.table = 120, split.cells  = c(20,20,20))
```

>>>>

#who_rec_low2_cp <- condprob(xX = nla_dat$log_chla,xY = nla_dat$log_micro, 
#                     xImpair = log10(mc_guides[5]), ProbComp = "gt", Exceed = "gte",
#                     ci = TRUE, R = 100, xW=wgt)

#mc_chla[5] <- get_chla(who_rec_low2_cp,0.5)

#who_rec_med1_cp <- condprob(xX = nla_dat$log_chla,xY = nla_dat$log_micro, 
#                     xImpair = log10(mc_guides[6]), ProbComp = "gt", Exceed = "gte",
#                     ci = TRUE, R = 100, xW=wgt)

#mc_chla[6] <- get_chla(who_rec_med1_cp,0.5)

#who_rec_med2_cp <- condprob(xX = nla_dat$log_chla,xY = nla_dat$log_micro, 
#                     xImpair = log10(mc_guides[7]), ProbComp = "gt", Exceed = "gte",
#                     ci = TRUE, R = 100, xW=wgt)

#mc_chla[7] <- get_chla(who_rec_med2_cp,0.5)
```

@@

#Methods

##Data

##Conditional Probability Analysis
We used a conditional probability analysis approach to explore associations between chlorophyll *a* concentrations  and World Health Organization (WHO) and U.S. Environmental Protection Agency (U.S. EPA) microcystin-LR health advisory levels [REFS].  Many levels have been suggested [Table \ref{tab:microcystin_levels}], but lakes with higher microcystin-LR concentrations in the NLA were rare. Only `r round(100*(length(nla_dat$MCYST_TL_UGL[nla_dat$MCYST_TL_UGL>=10])/nrow(nla_dat)),2)` % of lakes sampled had a concentration greater than 10.  Thus, for this analysis we focus on the microcystin concentrations that are better represented in the NLA data. These were  0.3, 1, 1.6, and 2 $\mu$g/L.

@@

(Figure \ref{fig:chla_micro_scatter})

(Figure \ref{fig:chla_micro_cdf})

(Figure \ref{fig:multi_cp_plot})

@@

```{r chla_micro_scatter, echo=FALSE, fig.width=8, fig.cap="Scatterplot showing association betweeen chlorophyll \\it{a} and microcystin-LR. \\label{fig:chla_micro_scatter}"}
plot_scatter(nla_dat,CHLA,MCYST_TL_UGL,viridis(1))
```

@@

```{r chla_micro_cdf, echo=FALSE, fig.width=8, fig.cap="Conditional probability plots showing association between the probability of exceeding various microcystin-LR (MLR) health advisory Levels. \\label{fig:chla_micro_cdf}"}
plot_cdf()
```

@@

```{r epa_child_cp_plot, echo=FALSE, fig.width=8, fig.cap="Conditional probability plots showing association between the probability of exceeding various microcystin-LR (MLR) health advisory Levels. \\label{fig:multi_cp_plot}"}
multiplot(
  plot_cp(epa_child_cp,x = "Log 10 Chlorophyll", 
          y = "Probability of Exceeding 0.3 ug/L MLR",
          title="U.S. EPA Children's Drinking Water Advisory Level"),
  plot_cp(who_drink_cp, x = "Log 10 Chlorophyll", 
          y = "Probability of Exceeding 1 ug/L MLR", 
          title="WHO Drinking Water Advisory Level"),
  plot_cp(epa_adult_cp, x = "Log 10 Chlorophyll", 
          y = "Probability of Exceeding 1.6 ug/L MLR", 
          title="U.S. EPA Adult Drinking Water Advisory Level"),
  plot_cp(who_rec_low1_cp, x = "Log 10 Chlorophyll", 
          y = "Probability of Exceeding 2 ug/L MLR", 
          title="WHO Recreational Water Advisory Level"),
  cols=2
)
```

@@

```{r microcystin_levels,echo=FALSE,results='asis', fig.cap="Various suggested microcystin-LR health advisory concentrations. \\label{fig:microcystin_levels}"}
source_of_stnd <- c("WHO",
                    "WHO",
                    "WHO",
                    "WHO",
                    "WHO",
                    "U.S. EPA",
                    "U.S. EPA")
type_of_stnd <- c("Recreational",
                  "Recreational",
                  "Recreational",
                  "Recreational",
                  "Drinking",
                  "Drinking",
                  "Drinking")
concentration <- c("2-4 $\\mu$g/L",
                   "10-20 $\\mu$g/L",
                   "20-2000 $\\mu$g/L",
                   ">2000 $\\mu$g/L",
                   "1 $\\mu$g/L",
                   "0.3 $\\mu$g/L",
                   "1.6 $\\mu$g/L")
#note <- c("Low probability of adverse health effects",
#          "Moderate probability of adverse health effects",
#          "High probability of adverse health effects, These higher #concentrations typically only seen in visible scums/algal mats",
#          "Very high probability of adverse health effects",
#          "",
#          "Children younger than school age",
#          "Individuals School age and older")
mc_conc_table <- data.frame(source_of_stnd, type_of_stnd, concentration) %>%
  arrange(type_of_stnd)
names(mc_conc_table)<-c("Source","Type","Concentration")
pandoc.table(mc_conc_table,justify = "left" ,split.table = 120, split.cells  = c(20,20,20))
```
